### 2022-11-24 dissertation/chapters/05-probability.tex
✨ (final-review): Up to chapter 4

\begin{quotation}
      \noindent ``Probability theory is nothing but common sense reduced to calculation.''
\end{quotation}
\begin{flushright}
      - Pierre-Simon Laplace
\end{flushright}

\noindent
Probability theory and statistics can be traced back to as early as the 18th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely seen as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of \index{Bayes' Theorem}Bayes' Theorem was introduced.

@@

Chapter \ref{chap:hhs} provided the reader with the concept of a \acf{HH}. This dissertation aims to develop a selection \acs{HH} that makes use of probability theory to select the best \acs{HH} to solve a given problem. This chapter aims to provide the reader with the necessary background information on probability theory and statistics. These are big fields and focus is put on the elements that are required to formulate the proposed \acf{BHH}. The remainder of the chapter is structured as follows.

@@

\noindent
In Equation~\eqref{eq:probability:overview:conditional} above, $P(B) > 0$. Note that conditional probability does not suggest causation. If an event $A$ has a high probability of occurring after observing event $B$, it does not necessarily mean that $A$ is caused by $B$. Conditional probability simply expresses the dependence amongst events.

@@

      \noindent
      otherwise, events $A$ and $B$ are said to be \textit{dependent} random events.
\end{definition}

@@

      \noindent
      If $A$ and $B$ are independent, then

@@

      \noindent
      Then the collection of sets $\{B_{1}$, $B_{2}$, $\dots$, $B_{K}\}$ is said to be a partition of S, the union of mutually exclusive subsets.
\end{definition}
\vspace*{0.5cm}

@@

\noindent
One of the many applications of \index{Bayes' Theorem}Bayes' Theorem is to do statistical inference. \index{Bayes' Theorem}Bayes' Theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence.

@@

\noindent
From Equation~\eqref{eq:probability:probability_distributions:beta:pdf} above, the normalising constant $B(\alpha, \beta)$, is defined below in Equation~\eqref{eq:probability:probability_distributions:beta:norm_const}.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:beta:norm_const}, $\Gamma$ is the Gamma function as defined in Equation~\eqref{eq:probability:probability_distributions:beta:gamma_func} below.

@@

\noindent
It should be noted that the Gamma function can also be written as shown in Equation~\eqref{eq:probability:probability_distributions:beta:gamma_func_alt} below.

@@

\noindent
The control parameters $\alpha$ and $\beta$, determine the shape of the distribution. There exists a special case where $\alpha = \beta$. This is referred to as the \textit{symmetric} \index{Beta probability distribution}Beta probability distribution. In the case where $\alpha = \beta = 1$ the distribution is equivalent to the uniform distribution over all points in its support. The \index{Beta probability distribution}Beta probability distribution for various values of $\alpha$ and $\beta$, including the symmetric version is presented in Figure \ref{fig:probability:probability_distributions:beta}.

@@

\noindent
The expected value of $x$ is given below in Equation~\eqref{eq:probability:probability_distributions:beta:expected_value}.

@@

\noindent
Similarly, the expected value of the natural logarithm of $x$ can be calculated as is shown below in Equation~\eqref{eq:probability:probability_distributions:beta:expected_value_ln}.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:beta:expected_value_ln} above, $\psi$ is the logarithmic derivative of the Gamma function, called the \textit{Digamma} function. The Digamma function is given below in Equation~\eqref{eq:probability:probability_distributions:beta:digamma}.

@@

\noindent
Finally, the mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:beta:mode} below.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:dirichlet:pdf} above, the normalising constant $B(\alpha)$, is defined in Equation~\eqref{eq:probability:probability_distributions:dirichlet:norm_cost}.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:dirichlet:pdf} above, $\alpha_{0}$ is defined in Equation~\eqref{eq:probability:probability_distributions:dirichlet:alpha_0} below.

@@

\noindent
Importantly, the set $\{x_{k}\}_{k=1}^{K}$ belongs to the standard $K-1$ probability simplex $S$, meaning that $x_{K} = 1 - \sum_{k=1}^{K-1}x_{k}$ with support $\forall_{k=1}^{K} x \in [0,1]$. Under the simplex $S$, this means that the sum over all values of the vector $\vec{x}$ must be 1. The simplex can thus be rewritten as $\sum_{k=1}^{K}x_{k} = 1$.

@@

\noindent
The expected value of $x$ can be calculated as shown in Equation~\eqref{eq:probability:probability_distributions:dirichlet:expected_value} below.

@@

\noindent
Similarly, the expected value of the natural logarithm of $x_{k}$ can be calculated as defined in Equation~\eqref{eq:probability:probability_distributions:dirichlet:expected_value_ln} below.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:dirichlet:expected_value_ln} above, $\psi$ is the Digamma function as defined in Equation~\eqref{eq:probability:probability_distributions:beta:digamma}. Finally, the mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:dirichlet:mode} below.

@@

\noindent
Equation~\eqref{eq:probability:probability_distributions:bernoulli:pmf} above, can also be expressed as show in Equation~\eqref{eq:probability:probability_distributions:bernoulli:alt} below.

@@

\noindent
The mean of the \index{Bernoulli probability distribution}Bernoulli probability distribution approaches $\theta$ over many samples according to the \acs{CLT}~\cite{ref:grinstead:1997}. Figure \ref{fig:probability:probability_distributions:bernoulli:coin} shows a simulation of a fair-coin flipping simulation. The mean converges to 0.5 for various sample sizes.

@@

\noindent
The expected value of the distribution is thus given according to Equation~\eqref{eq:probability:probability_distributions:bernoulli:expected_value} below.

@@

\noindent
The mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:bernoulli:mode} below.

@@

\noindent
Similar to the \index{Bernoulli probability distribution}Bernoulli probability distribution, the mean of the \index{Binomial probability distribution}Binomial probability distribution is $N\theta$ given the \acs{CLT} as was shown in Figure \ref{fig:probability:probability_distributions:bernoulli:coin}. The expected value of the \index{Binomial probability distribution}Binomial probability distribution is thus given in Equation~\eqref{eq:probability:probability_distributions:binomial:expected_value} below.

@@

\noindent
The mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:binomial:mode} below.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:categorical:pmf} above, $[x = k]$ is the \textit{Iversion Bracket}~\cite{ref:iverson:1962}, yielding 1 if $x = k$ and 0 otherwise. Given class $k$, the categorical distribution simply yields $\theta_{k}$ as is shown in Equation~\eqref{eq:probability:probability_distributions:categorical:pmf_k} below.

@@

\noindent
The random variable $x$ can also be encoded in binary format, yielding a vector $\vec{x} = (x_{1}, \dots, x_{K})$ of \index{Bernoulli probability distribution}Bernoulli probability distributions such that the support is $\forall_{k=1}^{K} x_{k} \in \{0, 1\}$. Importantly, if the outcome of the random event is of category $k$, then $x_{k} = 1$ and $\forall_{j=1}^{K} x_{j} = 0, j \neq k$ so that the standard $K-1$ probability simplex $S$ still holds. The \acs{PMF} of the \index{Categorical probability distribution}Categorical probability distribution can then be rewritten as follows in Equation~\eqref{eq:probability:probability_distributions:categorical:pmf_alt} below.

@@

\noindent
In Equation \ref{eq:probability:probability_distributions:categorical:pmf_alt} above, $\mathbbm{1}(x_{k})$ is the \textit{indicator function}, yielding 1 if $x_{k} = 1$ and 0 otherwise.

@@

\noindent
The support for the Multinomial is $\forall_{i=1}^{K} x_{k} \in \{1, \dots, N\}, \sum_{k=1}^{K}x_{k} = N$ and the \acs{PMF} for the \index{Multinomial probability distribution}Multinomial probability distribution is given below in Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf}.

@@

\noindent
Similar to the \index{Categorical probability distribution}Categorical probability distribution, the random variable $x$ can also be encoded in binary format, yielding an $N \times K$ matrix $X$ of \index{Bernoulli probability distribution}Bernoulli probability distributions. The support is then given as $X \in \{0, 1\}^{N \times K}, \forall_{i=1}^{N}\sum_{k=1}^{K} x_{i,k} = 1$ so that the standard $K-1$ probability simplex $S$ still holds for each trial. The \acs{PMF} of the \index{Multinomial probability distribution}Multinomial probability distribution can then be rewritten as is presented in Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt} below.

@@

\noindent
In Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt} above, $N_{k}$ is a summary variable, denoting the number of times a category $k$ occurs over all trials in $N$.

@@

\noindent
The Bernoulli likelihood is given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:likelihood} below.

@@

\noindent
By \index{Bayes' Theorem}Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:posterior} below.

@@

\noindent
Since the denominator sums to $1$, one could get rid of the denominator and constants for the Bernoulli likelihood and the Beta prior, by expressing the posterior as proportional to the likelihood times the prior as is shown in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:posterior_propto} below.

@@

\noindent
The posterior distribution has the same functional form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N_{1} + \alpha$ and $\beta' = N_{0} + \beta$. This shows that the \index{Beta probability distribution}Beta probability distribution is the conjugate prior used with the Bernoulli likelihood.

@@

\noindent
The likelihood of the Categorical and \index{Multinomial probability distribution}Multinomial probability distributions is given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:likelihood} below.

@@

\noindent
By \index{Bayes' Theorem}Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:posterior} below.

@@

\noindent
Since the denominator sums to $1$, one could get rid of the denominator and constants for the \index{Dirichlet probability distribution}Dirichlet prior, by expressing the posterior as proportional to the likelihood times the prior as is shown in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:posterior_propto} below.

@@

\noindent
The posterior distribution has the same form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N + \alpha$. This shows that the \index{Dirichlet probability distribution}Dirichlet probability distribution is the conjugate prior used with the Categorical/Multinomial likelihood.

@@

\noindent
If Bayesian statistics is used, the experimental design that was chosen does not matter. In Bayesian statistics it is common to use a \index{Beta probability distribution}Beta probability distribution as a prior distribution. If the prior distribution is sampled from $Beta(3,3)$ then using Bayesian analysis, the posterior distribution, according to the outcomes of this experiment, would yield $Beta(9,4)$.

@@

\noindent
It can be seen from Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} how the posterior distribution differs from the prior distribution. This is the nature of Bayesian analysis and is presented in more detail in the following section.

@@

\noindent
The marginal density or mass function of $Y_{1}, Y_{2}, \dots, Y_{N}$ is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:marginal_density} below.

@@

\noindent
Finally, the posterior density of $\theta \vert y_{1}, y_{2}, \dots, y_{n}$, according to \index{Bayes' Theorem}Bayes' Theorem, is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:posterior_density} below.

@@

\noindent
\citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} mentions that the posterior density summarises all the pertinent information about the parameter $\theta$ by making use of the information contained in the prior for $\theta$, as well as the information in the observed data/evidence.

@@

\noindent
In Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_1} above, $y_{i} = \{0,1\}$ and $0 < \theta < 1$. The likelihood $\mathcal{L}(y_{1}, y_{2}, \dots, y_{n} \vert \theta)$ is presented in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_2} below.

@@

\noindent
Then the joint likelihood of $Y_{1}, Y_{2}, \dots, Y_{N}, \theta$ from Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:joint_likelihood} can be formulated as in shown in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_3} below.

@@

\noindent
The marginal density of $Y_{1}, Y_{2}, \dots Y_{N}$ is then given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_4} below.

@@

\noindent
Finally, the posterior density of $\theta$ is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_5} below.

@@

\noindent
The posterior density has the same functional shape $\mathcal{A}(v)$ as the prior, yielding the update on prior parameters as $\alpha^{*} = \sum y_{i} + \alpha$ and $\beta^{*} = n - \sum y_{i} + \beta$.

>>>>

\begin{quotation}
      ``Probability theory is nothing but common sense reduced to calculation.''
\end{quotation}
\begin{flushright}
      - Pierre-Simon Laplace
\end{flushright}

Probability theory and statistics can be traced back to as early as the 18th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely seen as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of \index{Bayes' Theorem}Bayes' Theorem was introduced.

@@

Chapter \ref{chap:hhs} provided the reader with the concept of a \acf{HH}. This dissertation aims to develop a selection \acs{HH} that makes use of probability theory to select the best \acs{HH} to solve a given problem. This chapter aims to provide the reader with the necessary background information on probability theory and statistics. These are big fields and focus is put on the elements that are required to formulate the proposed \acf{BHH}. The remainder of the chapter is structured as follows:

@@

In Equation~\eqref{eq:probability:overview:conditional} above, $P(B) > 0$. Note that conditional probability does not suggest causation. If an event $A$ has a high probability of occurring after observing event $B$, it does not necessarily mean that $A$ is caused by $B$. Conditional probability simply expresses the dependence amongst events.

@@

      otherwise, events $A$ and $B$ are said to be \textit{dependent} random events.
\end{definition}

@@

      If $A$ and $B$ are independent, then

@@

      Then the collection of sets $\{B_{1}$, $B_{2}$, $\dots$, $B_{K}\}$ is said to be a partition of S, the union of mutually exclusive subsets.
\end{definition}
\vspace*{0.5cm}

@@

One of the many applications of \index{Bayes' Theorem}Bayes' Theorem is to do statistical inference. \index{Bayes' Theorem}Bayes' Theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence.

@@

From Equation~\eqref{eq:probability:probability_distributions:beta:pdf} above, the normalising constant $B(\alpha, \beta)$, is defined below in Equation~\eqref{eq:probability:probability_distributions:beta:norm_const}.

@@

In Equation~\eqref{eq:probability:probability_distributions:beta:norm_const}, $\Gamma$ is the Gamma function as defined in Equation~\eqref{eq:probability:probability_distributions:beta:gamma_func} below.

@@

It should be noted that the Gamma function can also be written as shown in Equation~\eqref{eq:probability:probability_distributions:beta:gamma_func_alt} below.

@@

The control parameters $\alpha$ and $\beta$, determine the shape of the distribution. There exists a special case where $\alpha = \beta$. This is referred to as the \textit{symmetric} \index{Beta probability distribution}Beta probability distribution. In the case where $\alpha = \beta = 1$ the distribution is equivalent to the uniform distribution over all points in its support. The \index{Beta probability distribution}Beta probability distribution for various values of $\alpha$ and $\beta$, including the symmetric version is presented in Figure \ref{fig:probability:probability_distributions:beta}.

@@

The expected value of $x$ is given below in Equation~\eqref{eq:probability:probability_distributions:beta:expected_value}.

@@

Similarly, the expected value of the natural logarithm of $x$ can be calculated as is shown below in Equation~\eqref{eq:probability:probability_distributions:beta:expected_value_ln}.

@@

In Equation~\eqref{eq:probability:probability_distributions:beta:expected_value_ln} above, $\psi$ is the logarithmic derivative of the Gamma function, called the \textit{Digamma} function. The Digamma function is given below in Equation~\eqref{eq:probability:probability_distributions:beta:digamma}.

@@

Finally, the mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:beta:mode} below.

@@

In Equation~\eqref{eq:probability:probability_distributions:dirichlet:pdf} above, the normalising constant $B(\alpha)$, is defined in Equation~\eqref{eq:probability:probability_distributions:dirichlet:norm_cost}.

@@

In Equation~\eqref{eq:probability:probability_distributions:dirichlet:pdf} above, $\alpha_{0}$ is defined in Equation~\eqref{eq:probability:probability_distributions:dirichlet:alpha_0} below.

@@

Importantly, the set $\{x_{k}\}_{k=1}^{K}$ belongs to the standard $K-1$ probability simplex $S$, meaning that $x_{K} = 1 - \sum_{k=1}^{K-1}x_{k}$ with support $\forall_{k=1}^{K} x \in [0,1]$. Under the simplex $S$, this means that the sum over all values of the vector $\vec{x}$ must be 1. The simplex can thus be rewritten as $\sum_{k=1}^{K}x_{k} = 1$.

@@

The expected value of $x$ can be calculated as shown in Equation~\eqref{eq:probability:probability_distributions:dirichlet:expected_value} below.

@@

Similarly, the expected value of the natural logarithm of $x_{k}$ can be calculated as defined in Equation~\eqref{eq:probability:probability_distributions:dirichlet:expected_value_ln} below.

@@

In Equation~\eqref{eq:probability:probability_distributions:dirichlet:expected_value_ln} above, $\psi$ is the Digamma function as defined in Equation~\eqref{eq:probability:probability_distributions:beta:digamma}. Finally, the mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:dirichlet:mode} below.

@@

Equation~\eqref{eq:probability:probability_distributions:bernoulli:pmf} above, can also be expressed as show in Equation~\eqref{eq:probability:probability_distributions:bernoulli:alt} below.

@@

The mean of the \index{Bernoulli probability distribution}Bernoulli probability distribution approaches $\theta$ over many samples according to the \acs{CLT}~\cite{ref:grinstead:1997}. Figure \ref{fig:probability:probability_distributions:bernoulli:coin} shows a simulation of a fair-coin flipping simulation. The mean converges to 0.5 for various sample sizes.

@@

The expected value of the distribution is thus given according to Equation~\eqref{eq:probability:probability_distributions:bernoulli:expected_value} below.

@@

The mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:bernoulli:mode} below.

@@

Similar to the \index{Bernoulli probability distribution}Bernoulli probability distribution, the mean of the \index{Binomial probability distribution}Binomial probability distribution is $N\theta$ given the \acs{CLT} as was shown in Figure \ref{fig:probability:probability_distributions:bernoulli:coin}. The expected value of the \index{Binomial probability distribution}Binomial probability distribution is thus given in Equation~\eqref{eq:probability:probability_distributions:binomial:expected_value} below.

@@

The mode of the distribution is given in Equation~\eqref{eq:probability:probability_distributions:binomial:mode} below.

@@

In Equation~\eqref{eq:probability:probability_distributions:categorical:pmf} above, $[x = k]$ is the \textit{Iversion Bracket}~\cite{ref:iverson:1962}, yielding 1 if $x = k$ and 0 otherwise. Given class $k$, the categorical distribution simply yields $\theta_{k}$ as is shown in Equation~\eqref{eq:probability:probability_distributions:categorical:pmf_k} below.

@@

The random variable $x$ can also be encoded in binary format, yielding a vector $\vec{x} = (x_{1}, \dots, x_{K})$ of \index{Bernoulli probability distribution}Bernoulli probability distributions such that the support is $\forall_{k=1}^{K} x_{k} \in \{0, 1\}$. Importantly, if the outcome of the random event is of category $k$, then $x_{k} = 1$ and $\forall_{j=1}^{K} x_{j} = 0, j \neq k$ so that the standard $K-1$ probability simplex $S$ still holds. The \acs{PMF} of the \index{Categorical probability distribution}Categorical probability distribution can then be rewritten as follows in Equation~\eqref{eq:probability:probability_distributions:categorical:pmf_alt} below.

@@

In Equation \ref{eq:probability:probability_distributions:categorical:pmf_alt} above, $\mathbbm{1}(x_{k})$ is the \textit{indicator function}, yielding 1 if $x_{k} = 1$ and 0 otherwise.

@@

The support for the Multinomial is $\forall_{i=1}^{K} x_{k} \in \{1, \dots, N\}, \sum_{k=1}^{K}x_{k} = N$ and the \acs{PMF} for the \index{Multinomial probability distribution}Multinomial probability distribution is given below in Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf}.

@@

Similar to the \index{Categorical probability distribution}Categorical probability distribution, the random variable $x$ can also be encoded in binary format, yielding an $N \times K$ matrix $X$ of \index{Bernoulli probability distribution}Bernoulli probability distributions. The support is then given as $X \in \{0, 1\}^{N \times K}, \forall_{i=1}^{N}\sum_{k=1}^{K} x_{i,k} = 1$ so that the standard $K-1$ probability simplex $S$ still holds for each trial. The \acs{PMF} of the \index{Multinomial probability distribution}Multinomial probability distribution can then be rewritten as is presented in Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt} below.

@@

In Equation~\eqref{eq:probability:probability_distributions:multinomial:pmf_alt} above, $N_{k}$ is a summary variable, denoting the number of times a category $k$ occurs over all trials in $N$.

@@

The Bernoulli likelihood is given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:likelihood} below.

@@

By \index{Bayes' Theorem}Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:posterior} below.

@@

Since the denominator sums to $1$, one could get rid of the denominator and constants for the Bernoulli likelihood and the Beta prior, by expressing the posterior as proportional to the likelihood times the prior as is shown in Equation~\eqref{eq:probability:conjugate_priors:binom_likelihood:posterior_propto} below.

@@

The posterior distribution has the same functional form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N_{1} + \alpha$ and $\beta' = N_{0} + \beta$. This shows that the \index{Beta probability distribution}Beta probability distribution is the conjugate prior used with the Bernoulli likelihood.

@@

The likelihood of the Categorical and \index{Multinomial probability distribution}Multinomial probability distributions is given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:likelihood} below.

@@

By \index{Bayes' Theorem}Bayes' Theorem, the posterior distribution with given prior data $\mathcal{D}$ is given in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:posterior} below.

@@

Since the denominator sums to $1$, one could get rid of the denominator and constants for the \index{Dirichlet probability distribution}Dirichlet prior, by expressing the posterior as proportional to the likelihood times the prior as is shown in Equation~\eqref{eq:probability:conjugate_priors:mult_likelihood:posterior_propto} below.

@@

The posterior distribution has the same form $\mathcal{A}(v)$ as the prior, but with updated prior parameters $\alpha' = N + \alpha$. This shows that the \index{Dirichlet probability distribution}Dirichlet probability distribution is the conjugate prior used with the Categorical/Multinomial likelihood.

@@

If Bayesian statistics is used, the experimental design that was chosen does not matter. In Bayesian statistics it is common to use a \index{Beta probability distribution}Beta probability distribution as a prior distribution. If the prior distribution is sampled from $Beta(3,3)$ then using Bayesian analysis, the posterior distribution, according to the outcomes of this experiment, would yield $Beta(9,4)$.

@@

It can be seen from Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} how the posterior distribution differs from the prior distribution. This is the nature of Bayesian analysis and is presented in more detail in the following section.

@@

The marginal density or mass function of $Y_{1}, Y_{2}, \dots, Y_{N}$ is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:marginal_density} below.

@@

Finally, the posterior density of $\theta \vert y_{1}, y_{2}, \dots, y_{n}$, according to \index{Bayes' Theorem}Bayes' Theorem, is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:posterior_density} below.

@@

\citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} mentions that the posterior density summarises all the pertinent information about the parameter $\theta$ by making use of the information contained in the prior for $\theta$, as well as the information in the observed data/evidence.

@@

In Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_1} above, $y_{i} = \{0,1\}$ and $0 < \theta < 1$. The likelihood $\mathcal{L}(y_{1}, y_{2}, \dots, y_{n} \vert \theta)$ is presented in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_2} below.

@@

Then the joint likelihood of $Y_{1}, Y_{2}, \dots, Y_{N}, \theta$ from Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:joint_likelihood} can be formulated as in shown in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_3} below.

@@

The marginal density of $Y_{1}, Y_{2}, \dots Y_{N}$ is then given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_4} below.

@@

Finally, the posterior density of $\theta$ is given in Equation~\eqref{eq:probability:bayesian_statistic:bayesian_analysis:step_5} below.

@@

The posterior density has the same functional shape $\mathcal{A}(v)$ as the prior, yielding the update on prior parameters as $\alpha^{*} = \sum y_{i} + \alpha$ and $\beta^{*} = n - \sum y_{i} + \beta$.
